Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2021-10-13 methods_paper/fabl-paper.tex
fabl paper ready for review

	Many statistical record linkage methods are extensions of seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}. Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match using a likelihood ratio test. Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{winkler1991application, fair2004generalized, wagner2014person, gill2003english}. Although the need to compute comparisons for all record pairs initially restricted this family of methods, recent work by \cite{enamorado2019using} used innovative hashing techniques to scale the approach to handle large datasets.

	The independent pairwise matching assumption from Fellegi and Sunter is popular mainly for its mathematical simplicity, but is often unreasonable in practice. In many situations, we know that there are no duplications within a database, meaning that one record from one database should be linked with at most one record in the other database. Here, additional declared matches are known by assumption to be false. Many extensions to \cite{fellegi_theory_1969} resolve these false matches as a post-processing step \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.

	Alternatively, one can embed one-to-one matching requirements into the model specification itself, at an additional computational cost. \cite{larsen2005} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space. \cite{fortunato_2010} used simulated annealing to target the space of permitted matches under the one-to-one constraint, but the method was so computationally intensive it could used to link databases of less than 100 records. Most recently, \cite{sadinle_bayesian_2017} proposed \emph{Beta Record Linkage} (BRL), using an innovative prior over the space of bipartite matchings to strictly enforce one-to-one reqiurements throughout his Gibbs sampler. Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeller can weight the relative importance of false positives and false negatives, and identify records pairings to be be decided through clerical review. Although it was shown to work on larger tasks than previous one-to-one methods, \texttt{BRL} becomes slow when working with larger datasets, and unusable when the number of pairwise record comparisons becomes too large.

	In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage. This allows us to employ hashing techniques that hasten calculations and reduce computational costs. Furthermore, we can compute the pairwise record comparisons over large datasets via parallel computing and reduce memory costs through our proposed method of \emph{storage efficient indexing}. Throughout, we argue that even in cases where a bipartite matching is desired, \textcolor{red}{our proposed approach and relaxation} (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed. Open source software to use \texttt{fabl} in \texttt{R} is available through Github.

	In what follows, Section \ref{sec:review-of_prior-work} reviews more thoroughly the Fellegi Sunter and Beta Record Linkage methods. Then, Section \ref{sec:fast-beta-linkage} provides the model specification under \texttt{fabl} and derives the Gibbs sampler for posterior inference. Next, Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake. Lastly, in Sections \ref{sec:simulations} and \ref{sec:case-studies} we use simulation studies and a case study of homicides from the El Salvadoran Civil War to show that our method has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}. We conclude in Section \ref{discussion} with a discussion of directions for future work.

@@

Consider two databases $\bm{X}_1$ and $\bm{X}_2$ with respective sizes $n_1$ and $n_2$. Without loss of generality, denote files such that $n_1 \geq n_2$. In the context of bipartite matching, we assume that there are duplications across, but not within, each database. Under this framework, the set of matches across datasets can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1 \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0 \quad \text{otherwise}.\\
	\end{cases}
\end{align}
Though intuitive, this sparse matrix representation can become costly for large linkage tasks or for long posterior samplers. More compactly, bipartite matching can also be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j \quad  \quad \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
	\end{cases}
\end{align}
Depending on which representation is the most convenient, we can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot)$ is the indicator function.

@@

\subsection{Comparison Data}

Intuitively, records that are co-referent or rather refer to the same entity should be similar; records that are non co-referent and refer to different entities should not be similar. \cite{fellegi_theory_1969} proposed encoding this is using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that 
$\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ In most cases, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets.

The simplest way to compare two records is check for agreement or disagreement, and this is commonly used for categorical variables. For more complex measurements, we can take into account partial agreement to more richly characterize the comparison; for numerical data, we can use absoluate difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance. This allows us to divide our records into a set of similarity values into different levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}. Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$. Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors in the following way: 
$$\gamma_{ij}^f = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$$

@@

\subsection{Fellegi-Sunter Model}
The comparison data $\bm{\gamma}_{ij}$ is not sufficient to determine whether a record is a match or nonmatch because of errors that naturally occur in the data. This motivated \cite{fellegi_theory_1969} to consider the following likelihood ratio
\begin{align}
	\label{eqn:wts}
	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}
\end{align}
as a weight to estimate if record pairs are a match or nonmatch. This ratio will be large if there is strong evidence of the pair being a match, and small otherwise.

In practice $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated. Under the Fellegi Sunter model, we assume that comparison vectors are independent given their matching status, and that match status of the record pairs are independent, allowing us to model 
the comparison data using mixture models in the following way:
\begin{align}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
\end{align}
where $p$ is the proportion of records that match, and $\bm{m}$ and $\bm{u}$ are vectors of parameters for the matching and nonmatching record pairs. These unknown parameters are often estimated using the EM algorithm.

After estimating $\bm{m}$ and $\bm{u}$, the Fellegi Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ declared as nonlinks (with those such that $T_u < w_{ij} < T_m$ left undetermined). This is a problem in practice as transitive closures are violated. Since each $w_{ij}$ is calculated independently, there are often situation in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.

@@

The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi Sunter weights among matched pairs. \cite{jaro1989} provided no theoretical justification for using this approach, however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, equation \ref{eqn:jaro} is the maximum likelihood estimate.

%One criticism of using mixture models as proposed by \cite{jaro1989} is the fact that the decision rule will lead to ``many-to-many assignments (breaking transitive closures). One way of resolving this issue to incorporate the constraint directly into the model \citep{Fortinietal01, Matsakis10, liseo_jos_2011, liseo_2011, Larsen05, Larsen12, gutman_bayesian_2013, sadinle_bayesian_2017}, which is appealing as one then obtains uncertainty quantification under the Bayesian paradigm, but often has to sacrifice scaling to very large databases. On the other hand, one can turn to post-processing methods in practical situations. In this paper, we attempt to bridge the best of both approaches. Specifically, we propose a model in the spirit of \cite{sadinle_bayesian_2017} and \cite{enamorado_using_2019}. We relax the one-to-one matching constraint proposed by  \cite{sadinle_bayesian_2017} for purely computational reasons, and resolve this in a post-processing step. In addition, we propose our own computational speeds up and those in the spirit of \cite{enamorado_using_2019}. Our framework is then embarassingly parallel, available in the \texttt{fabl} package, and evaluated on a real case study and simulation studies illustrating that any loss in accuracy due to removing the constraint is minimal.

\subsection{Beta Record Linkage Model}

One criticism of using mixture models is that the decision rule will lead to ``many-to-many assignments,'' breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows:
%our model can be written as a Bayesian model for comparison data. In general, it has the same form of \cite{sadinle_bayesian_2017} and can be written as follows:
\begin{align}
	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
	& \bm{Z} \sim \mathcal{B}, \notag
\end{align}
where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m}), \mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the \emph{beta distribution for bipartite matching}, given by
$$P(\mathbf{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\mathbf{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\mathbf{Z}) + \alpha_{\pi}, n_2 - \mathbf{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, and hyper parameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate
records when sampling \(Z_j\). This creates a dependency that makes the
sampler \emph{inherently serial}, which makes the sampler slow when working on large linkage tasks.

@@

In practice, it is not feasible to make all-to-all record comparisons as the computationally complexity is of the order $O(n_1 \times n_2).$ The most common solution is utilize blocking, which places similar records into partitions, or ``blocks'' to reduces this computational burden. In deterministic blocking, the modeller chooses a field thought to be highly reliable, and only compares records that agree on that field. The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains. However, blocking on an unreliable field can lead to many missed matches, making this form of blocking undesirable in many situations.

In this paper, we use hashing and new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without the need for blocking. This is useful when there is no reliable blocking field available, or one desires estimates of model parameters for the entire sample in question. In practice, the method the follows can be combined with blocking for even greater computational gains, but all derivations, simulations, and case studies are presented without blocking.

@@

Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and that each record $i \in \bm{X}_1$ is equally likely to be that match. $\pi$ itself is given a prior distribution with hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ to encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$.. One choice of uninformative prior might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that nonmatches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\mathbf{Z}\). We encourage the reader to integrate out $\pi$ from the $Z_j|\pi$ under these two settings to gain deeper intuition about this prior.

Note that linkage in this setting is conducted at the \emph{record} level, rather than at the \emph{record pair} level as in the Fellegi Sunter model. That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $\bm{X_2}$ that have matches, while $\lambda$ in Fellegi Sunter estimate that proportion of record pairs that have matches. We find $\pi$ to be more interpretable parameter than $\lambda$, especially considering that the overwhelming majority of pairs are non matches, meaning that $\lambda$ necessarily tends towards 0 as the size of the linkage task grows. Additionally, we emphasize that while the Fellegi Sunter model makes $n_1 \times n_2$ many independent matching decisions and \texttt{BRL} makes $n_2$ many dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_2$ many independent matching decisions. As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows us to achieve many of the accuracy gains from \texttt{BRL} while maintaining the efficiency gains possible by exploiting independence.

Importantly, this independence means that we have relaxed one-to-one requirement from \texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). This behavior may be desirable in its own right; \cite{Newcombe1962b} was interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable. In situations when a bipartite estimate is desired, we can use a simple post processing procedure shown in \ref{bayes-estimate} to resolve illegal matchings.

@@

\begin{align*}
	p(\mathbf{Z}, \Phi, \pi|\Gamma) &\propto \mathcal{L}(\bm{Z}, \Phi \mid \Gamma^{obs}) p(\mathbf{Z} | \pi) p(\Phi) p(\pi) \\
	&\propto \prod_{j=1}^{n_2}  \prod_{i=1}^{n_1}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
	&\times  \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{\alpha_{fl} - 1}  \times\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\beta_{fl} - 1} \\
	&\times \prod_{j=1}^{n_2} \left[I(Z_j \leq n_1)\frac{1}{n_1}\pi + I(Z_j > n_1)(1 - \pi)\right] \\
	&\times \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1}
\end{align*}
This factorization leads to following Gibbs Sampler:

\underline{Sample $\Phi^{(s+1)}|\Gamma, \mathbf{Z}^{(s)}$:}
The \(\mathbf{m}\) and \(\mathbf{u}\) parameters are updated through
standard multinomial-dirichlet mechanics. Thus we have

\[\mathbf{m}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\mathbf{Z}), \ldots, \alpha_{fL_f}(\mathbf{Z}))\]
\[\mathbf{u}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\mathbf{Z}), \ldots, \beta_{fL_f}(\mathbf{Z}))\]
where
\(\alpha_{fl}(\mathbf{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j = i)\)
and
\(\beta_{fl}(\mathbf{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j \neq i)\).

>>>>

	Many statistical record linkage methods are extensions of the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}. Specifically, Fellegi and Sunter created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match using a likelihood ratio test. Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{winkler1991application, fair2004generalized, wagner2014person, gill2003english}. Although the need to compute comparisons for all record pairs initially restricted this family of methods, recent work by \cite{enamorado2019using} used innovative hashing techniques to scale the approach to handle large datasets.

	The independent pairwise matching assumption from Fellegi and Sunter is popular mainly for its mathematical simplicity, but is often unreasonable in practice. In many situations, we know that there are no duplications within a database, meaning that one record from one database should be linked with at most one record from the other. Here, additional declared matches are known by assumption to be false. Many extensions to \cite{fellegi_theory_1969} resolve these false matches as a post-processing step \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.

	Alternatively, one can embed one-to-one matching requirements into the model specification itself, at an additional computational cost. \cite{larsen2005} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space. \cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method was so computationally intensive it could only be used to link databases of less than 100 records. Most recently, \cite{sadinle_bayesian_2017} proposed \emph{Beta Record Linkage} (\texttt{BRL}), using an innovative prior over the space of bipartite matchings to strictly enforce one-to-one reqiurements throughout his Gibbs sampler. Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeller can weight the relative importance of false positives and false negatives, and identify records pairings to be be decided through clerical review. Although it was shown to work on larger tasks than previous one-to-one methods, \texttt{BRL} becomes slow when working with larger datasets, and unusable when the number of pairwise record comparisons becomes too large.

	In this paper, we propose \emph{fast beta linkage} (\texttt{fabl}), which extends the the Beta Record Linkage model for increased efficiency and scalability. Specifically, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage. This allows us to (1) employ hashing techniques that hasten calculations and reduce computational costs; (2) compute the pairwise record comparisons over large datasets via parallel computing; and (3) reduce memory costs through our proposed method of \emph{storage efficient indexing}. Throughout, we argue that even in cases where a bipartite matching is desired, our proposed approach (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed. Open source software to use \texttt{fabl} in \texttt{R} is available through Github.

	In what follows, Section \ref{sec:review-of_prior-work} reviews more thoroughly the Fellegi Sunter and Beta Record Linkage methods. Then, Section \ref{sec:fast-beta-linkage} provides the model specification under \texttt{fabl} and derives the Gibbs sampler for posterior inference. Next, Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake. Lastly, Sections \ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and a case study of homicides from the El Salvadoran Civil War. Section \ref{discussion} concludes our paper with a discussion of directions for future work.

@@

Consider two databases $\bm{X}_1$ and $\bm{X}_2$ with respective sizes $n_1$ and $n_2$. Without loss of generality, denote the files such that $n_1 \geq n_2$. In the context of bipartite matching, we assume that there are duplications across, but not within, each database. Under this framework, the set of matches across datasets can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1 \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0 \quad \text{otherwise}.\\
	\end{cases}
\end{align}
Though intuitive, this sparse matrix representation can become costly for large linkage tasks or for long posterior samplers. More compactly, bipartite matching can also be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j \quad  \quad \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
	\end{cases}
\end{align}
Depending on which representation is the most convenient, we can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot)$ is the indicator function.

@@

\subsection{Comparison Data}
\label{comparison-data}

Intuitively, co-referent records (those that refer to the same entity) should be similar; records that are not co-referent should not be similar. \cite{fellegi_theory_1969} proposed encoding this is using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that 
$\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ In most cases, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets.

The simplest way to compare two records is to check for agreement or disagreement, and this is commonly used for categorical variables. For more complex measurements, we can take into account partial agreement to more richly characterize the comparison; for numerical data, we can use absoluate difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance. We can then propose thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}. Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$. Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors in the following way: 
$$\gamma_{ij}^f = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$$

@@

\subsection{Fellegi-Sunter Model}
\label{fellegi-sunter}
The comparison data $\bm{\gamma}_{ij}$ is not sufficient to determine whether a record is a match or nonmatch because of errors that naturally occur in the data. This motivated \cite{fellegi_theory_1969} to consider the following likelihood ratio
\begin{align}
	\label{eqn:wts}
	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}
\end{align}
as a weight to estimate if record pairs are a match or nonmatch. This ratio will be large if there is strong evidence of the pair being a match, and small otherwise.

In practice $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated. Under the Fellegi Sunter model, we assume both that comparison vectors are independent given their matching status, and that the match status of the record pairs are independent. This allows us to model 
the comparison data using mixture models in the following way:
\begin{align}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
\end{align}
where $p$ is the proportion of records that match, and $\bm{m}$ and $\bm{u}$ are vectors of parameters for the matching and nonmatching record pairs. These unknown parameters are often estimated using the EM algorithm.

After estimating $\bm{m}$ and $\bm{u}$, the Fellegi Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as nonlinks (with those such that $T_u < w_{ij} < T_m$ left undetermined). This is a problem in practice as transitive closures are violated. Since each $w_{ij}$ is calculated independently, there are often situation in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.

@@

The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi Sunter weights among matched pairs. Jaro provided no theoretical justification for using this approach, however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, equation \ref{eqn:jaro} is the maximum likelihood estimate for bipartite matching.

\subsection{Beta Record Linkage Model}
\label{BRL}

One criticism of using mixture models is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows:
\begin{align}
	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
	& \bm{Z} \sim \mathcal{B}, \notag
\end{align}
where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the \emph{beta distribution for bipartite matching}, given by
$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - \bm{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, and hyper parameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate
records when sampling \(Z_j\). This creates a dependency that makes the
sampler \emph{inherently serial}, which makes the sampler slow when working on large linkage tasks.

@@

In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden. In deterministic blocking, the modeller chooses a field thought to be highly reliable, and only compares records that agree on that field. The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains. However, blocking on an unreliable field can lead to many missed matches, making this form of blocking undesirable in many situations.

In this paper, we use hashing and new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without the need for blocking. This is useful when there is no reliable blocking field available, or one desires estimates of model parameters for the entire sample in question. In practice, the following method can be combined with blocking for even greater computational gains, but all derivations, simulations, and case studies are presented without blocking.

@@

Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and that each record $i \in \bm{X}_1$ is equally likely to be that match. $\pi$ itself is given a prior distribution with hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ to encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$. One choice of uninformative prior might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that nonmatches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\). We encourage the reader to integrate out $\pi$ from $Z_j|\pi$ under these two settings to gain deeper intuition about this prior.

Note that linkage in this setting is conducted at the \emph{record} level, rather than at the \emph{record pair} level as in the Fellegi Sunter model. That is, $\pi$ under \texttt{fabl} estimates the proportion of \emph{records} in $\bm{X_2}$ that have matches, while $\lambda$ in Fellegi Sunter estimates the proportion of \emph{record pairs} that are matches. We find $\pi$ to be more interpretable parameter than $\lambda$; especcially in the bipartite case, there are at most $n_2$ matching pairs out of $n_1 n_2$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_1}$ and tends towards 0 as the size of the linkage task grows.  necessarily tends towards 0 as the size of the linkage task grows. Additionally, we emphasize that while the Fellegi Sunter model makes $n_1 \times n_2$ many independent matching decisions and \texttt{BRL} makes $n_2$ many dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_2$ many independent matching decisions. As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows us to achieve many of the accuracy gains from \texttt{BRL} while maintaining the efficiency gains possible by exploiting independence.

Importantly, this independence means that we have relaxed the one-to-one requirement from \texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). This behavior may be desirable in its own right; \cite{Newcombe1962b} were interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable. In situations when a bipartite estimate is desired, we can use a simple post processing procedure shown in \ref{bayes-estimate} to resolve illegal matchings.

@@

\begin{align*}
	p(\bm{Z}, \Phi, \pi|\Gamma) &\propto \mathcal{L}(\bm{Z}, \Phi \mid \Gamma^{obs}) p(\bm{Z} | \pi) p(\Phi) p(\pi) \\
	&\propto \prod_{j=1}^{n_2}  \prod_{i=1}^{n_1}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
	&\times  \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{\alpha_{fl} - 1}  \times\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\beta_{fl} - 1} \\
	&\times \prod_{j=1}^{n_2} \left[I(Z_j \leq n_1)\frac{1}{n_1}\pi + I(Z_j > n_1)(1 - \pi)\right] \\
	&\times \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1}
\end{align*}
This factorization leads to following Gibbs Sampler:

\underline{Sample $\Phi^{(s+1)}|\Gamma, \bm{Z}^{(s)}$:}

The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-dirichlet mechanics. Thus we have

$$\bm{m}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z}))$$
$$\bm{u}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z}))$$
where
\(\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j = i)\)
and
\(\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j \neq i)\).

\underline{Sample $\pi^{(s+1)}|\bm{Z}^{(s)}$:} As a function of
\(\pi\), the linkage structure parameter \(\bm{Z}\) is sequence
of successes (when \(z_j < n_A + 1\)) and failures (when
\(z_j = n_A + 1\)), and therefore
\(p(\bm{Z}|\pi) = \mathcal{L}(\pi|\bm{Z})\) is
determined only by the number of duplicates
$n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by
$\bm{Z}$. Thus we have
